---
title: "Medicin Consumption in Portugal"
author: "Martim Almeida"
date: "2/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F)

library(data.table)
library(here)
library(ggplot2)
library(kableExtra)

```

```{r, data_work, warning=F}

data <- fread(here('data', 'HEALTH_PHMC_25022020175027742.csv'))

portugal <- data[Country == 'Portugal']
spain <- data[Country == 'Spain']
finland <- data[Country == 'Finland']



portugal[Measure == '% of total sales' & Year == 2018, .(Variable, Value, Year)][order(-Value)]
portugal[Measure == '% of total sales' & Year == 2011, .(Variable, Value, Year)][order(-Value)]

spain[Measure == '% of total sales' & Year == 2017, .(Variable, Value, Year)][order(-Value)]
spain[Measure == '% of total sales' & Year == 2011, .(Variable, Value, Year)][order(-Value)]

finland[Measure == '% of total sales' & Year == 2017, .(Variable, Value, Year)][order(-Value)]
finland[Measure == '% of total sales' & Year == 2011, .(Variable, Value, Year)][order(-Value)]

data[Measure == '% of total sales' & Variable == 'N05B-Anxiolytics' & Year == 2017, .(Variable, Country, Value )][order(-Value)]
data[Measure == '% of total sales' & Variable == 'N05B-Anxiolytics' & Year == 2011, .(Variable, Country, Value )][order(-Value)]

data[Measure == '% of total sales' & Variable == "A02A-Antacids" & Year == 2017, .(Country, Value )][order(-Value)]

data[Measure == '% of total sales' & Variable == "N06A-Antidepressants"  & Year == 2017, .(Country, Value )][order(-Value)]


m1 <- "Defined daily dosage per 1 000 inhabitants per day"

v5 <- "A02A-Antacids"  
  
v2 <- "C07-Beta blocking agents"  

v4 <- "N06A-Antidepressants"

v3 <- "G03-Sex hormones and modulators of the genital system" 

v1 <- 'N05B-Anxiolytics'


data[Measure == m1 & Variable == 'N05B-Anxiolytics' & Year == 2017, .(Variable, Country, Value )][order(-Value)]
data[Measure == m1 & Variable == 'N05B-Anxiolytics' & Year == 2011, .(Variable, Country, Value )][order(-Value)]

data[Measure == m1 & Variable == "A02A-Antacids" & Year == 2017, .(Country, Value )][order(-Value)]

data[Measure == m1 & Variable == "N06A-Antidepressants"  & Year == 2017, .(Country, Value )][order(-Value)]

data[Measure == m1 & Variable == v2 & Year == 2017, .(Country, Value )][order(-Value)]

data[Measure == m1 & Variable == v3 & Year == 2017, .(Country, Value )][order(-Value)]

data[Measure == m1   & Variable == "N-Nervous system" & Year == 2017, .(Country, Value )][order(-Value)]
```

# Medicine Consumption
## Introduction
Increase in medicine consumption could be a good indicator of health status on a population. It can show trends in some diseases that do not express in mortality rates but still affect quality of life of individuals. It is also a motive of concern due to chronical diseases increase and the need to treat them with pharmacological aids. OCDE gathered data from 2011 to 2017/2018 (depends on the country) of medicine consumption, measuring the following variables:
  
*"Defined daily dosage per 1 000 inhabitants per day"  
*"Million of national currency units"    
*"Million US$ at exchange rate"   
*"Million US$, purchasing power parity"  
*"/capita, US$ exchange rate"    
*"/capita, US$ purchasing power parity"   
*"% of total sales"   
*"% share of generics (value)"   
*"% share of generics (volume)"  
  
This report is a basic exploratory analysis of the data avaiable by OCDE, with focus on portugal, comparing with other countries and showing the trends along the time.

## Types of pharmacological drugs

```{r}

variables <- unique(data[, Variable])
table_1 <- data.table(
                      'Drug classification' = variables[1:19],
                      ' ' = variables[20:28])

kable(table_1) %>%
  kable_styling(bootstrap_options = 'striped')


```



## Exploratory Analysis

```{r}
ggplot(data[Variable == v1 & Measure == m1], aes(Year, Value)) +
  geom_line() +
  facet_wrap(~Country)
```




